\section{$O\br{\log n\log\log n}$-approximation for \textsc{HAC}}

Below we show how to devise an approximation algorithm for \textsc{HAC} consisiting of two steps. In the preprocessing step we use mathemathical programming in order to build a guide for our next algorithmic step. To do so, we start by encoding the problem using an \textsc{ILP}. We merge the properties of the \textsc{ILP} for \textsc{FES} used by \cite{ApproximatingMinimumFeedbackSetsAndMulticutsInDirectedGraphs} with the \textsc{SDP} for \textsc{HC} used by Charikar et. al in \cite{ApproximateHierarchicalClusteringViaSparsestCutAndSpreadingMetrics}.

Our relaxation of this \textsc{ILP} is twofold and each step serves two purposes, either choosing the subset of vertices, whose cycle we need to cut at each level or hoosing the edges which need to be cut in order to decrease the size of the instance efficiently. To achieve the former of these goals, we relax the integrality of the variables to get an \textsc{LP} relaxation which can be solved in polynomial time using the ellipsoid method. Then, we do a partial rounding routine to get an integral solution for the variables responsible for choosing the subset of vertices, whose cycle we need to cut at each level. By doing so, we obtain an partially integral solution to the \textsc{LP} relaxation without excessive increase in cost. This relaxation can be then decomposed into several instances of the \textsc{FES-LP} relaxation which will come useful during the rercursive processing. In the second step of preprocessing, we remove the variables reponsible for the above choice and we relax the remaining variables using the $\ell_2^2$-metric constraints to get an \textsc{SDP} relaxation. This \textsc{SDP} relaxation can be solved in polynomial time using interior point methods. Similarly, as before we decompose this \textsc{SDP} into several instances of the \textsc{SDP} relaxation for \textsc{$k$-BP} which we will use during the recursive step.

During the recursive step, we use the above decomposition of both \textsc{LP} and \textsc{SDP} relaxations to guide our recursive algorithm. At each level of the recursion, we pick the level $t=n\br{G}/4$ and we use the solution for the \textsc{LP} at $t$ as a valid solution for the \textsc{FES} problem. Therefore, we can use the rounding algorithm by \cite{ApproximatingMinimumFeedbackSetsAndMulticutsInDirectedGraphs} to find a subset of edges $S$ to cut at level $t$ such that the cost of cutting these edges is comparable to the cost of the \textsc{LP} solution at level $t$. After cutting these edges, we get several connected components. Let $H$ be such a component. If $H$ is a tree, we are done with this component. If $n\br{H}\leq n\br{G}/2$, we can process $H$ recursively. Otherwise, we use the solution for the \textsc{SDP} at level $t$ as a valid solution for the \textsc{$k$-BP} with $k=4$. Therefore, we can use the rounding algorithm by \cite{PartitioningGraphsIntoBalancedComponents} to find a subset of edges $S'$ to cut such that the size of all connected components after cutting $S'$ is at most $n\br{G}/2$ and the cost of cutting these edges is comparable to the cost of the \textsc{SDP} solution at level $t$. After cutting these edges, we get several connected components which we process recursively.

Since we decompose the \textsc{LP} into separate levels by summing up over all the levels of the recursion we will be able to show that our algorithm achieves an approximation ratio of $O\br{\alpha\br{n}+\beta\br{n}}$, where $\alpha\br{n}$ is the approximation ratio of the rounding algorithm for \textsc{$k$-BP} and $\beta\br{n}$ is the approximation ratio of the rounding algorithm for \textsc{FES}. Since $\alpha\br{n}=O\br{\log n\log\log n}$ and $\beta\br{n}=O\br{\sqrt{\log n}}$, we get an approximation ratio of $O\br{\log n\log\log n}$ for \textsc{HAC}.

\subsection{\textsc{ILP} formulation of \textsc{HAC} and its two relaxations}
    We decompose the clustering tree $T$ into $n$ separate edge cut levels. For each level $t$, we require one of the two properties to hold. Each cluster is either of size at most $t$ or it is a tree. The partition at level $t-1$ is a superset of partition at level $t$. For each pair of vertices $i,j\in V$ and level $1\in [n]$ we define a variable $x_{ij}^t$ to be $1$ if $i$ and $j$ are in different clusters at level $t$ and $0$ otherwise. We also define a variable $y_i^t$ to be $1$ if all cycles in $G$ containing $i$ are cut at level $t$ and $0$ otherwise. We have the following constraints:
    \begin{enumerate}
        \item \textbf{refinement:} $x_{ij}^t \leq x_{ij}^{t-1}$ for all $i,j\in V$ and $t\in [n]$. This is because if vertices $i$ and $j$ are in different clusters at level $t$, they must also be in different clusters at level $t-1$. Moreover, $y_i^t \leq y_i^{t-1}$ for all $i\in V$ and $t\in [n]$. This is because if all cycles containing vertex $i$ are cut at level $t$, they must also be cut at level $t-1$.
        \item \textbf{triangle inequality:} $x_{ik}^t \leq x_{ij}^t + x_{jk}^t$ for all $i,j,k\in V$ and $t\in [n]$. This is because if vertices $i$ and $k$ are in the same cluster at level $t$, then vertices $i$ and $j$ must be in the same cluster or vertices $j$ and $k$ must be in the same cluster at level $t$.
        \item \textbf{$\ell_2^2$-metric:} $x_{ij}^t = \frac{1}{2}\|v_i^t - v_j^t\|_2^2$ for all $i,j\in V$ and $t\in [n]$. The triangle inequality is implied by the $\ell_2^2$-metric constraint. With each vertex $i$ we associate a unit vector $v_i^t$ in a high-dimensional space. All vertices in one cluster at level $t$ are mapped to the same vector, while vertices in different clusters are mapped to orthogonal vectors. For completeness we also include $y_{i}^t = \frac{1}{2}\|u_{i,1}^t - u_{i,2}^t\|_2^2$. Note that this constraint is merely a vehicle to simulate $y_i^t\in[0,1]$ as a semidefinite constraint.
        \item \textbf{cycle cutting:} For every cycle $C$ in $G$, every vertex $i\in C$ and every level $t\in [n]$ we have $\sum_{ij\in C} x_{ij}^t \geq y_{i}^t$. This is because if all cycles containing vertex $i$ are cut at level $t$, then at least one edge in every cycle containing vertex $i$ must be cut at level $t$. Additionally, we have $y_i^1=1$, since every cycle eventually needs to be cut.
        \item \textbf{spreading:} For every vertex $i\in V$ and every level $t\in [n]$ we have $\sum_{j\in V} x_{ij}^t \geq \br{1-y_i^t}\cdot (n-t)$. This is because if the cluster containing vertex $i$ at level $t$ has size at most $t$, then either all cycles containing vertex $i$ are cut at level $t$ or at least $n-t$ vertices are in different clusters than vertex $i$ at level $t$.
    \end{enumerate}

    Our relaxation is two-fold. Firstly, we introduce the following \textsc{ILP} formulation \textsc{ILP-HAC} for the \textsc{HAC} problem:
    \begin{align}
        \min \quad &\sum_{t=1}^n \sum_{ij\in E} c\br{i,j} \cdot x_{ij}^t\\
        \text{subject to} \quad &x_{ij}^t \leq x_{ij}^{t-1} \quad \forall i,j\in V, t\in [n],\\
        &y_i^t \leq y_i^{t-1} \quad \forall i\in V, t\in [n],\\
        &y_i^1 = 1 \quad \forall i\in V,\\
        &x_{ik}^t \leq x_{ij}^t + x_{jk}^t \quad \forall i,j,k\in V, t\in [n],\\
        &\sum_{ij\in C} x_{ij}^t \geq y_{i}^t \quad \forall \text{ cycles } C, i\in C, t\in [n],\\ \label{cycle_constraint_ILP}
        &\sum_{j\in V} x_{ij}^t \geq \br{1-y_i^t}\cdot (n-t) \quad \forall i\in V, t\in [n],\\ \label{spreading_constraint_ILP}
        &x_{ij}^t \in \{0,1\} \quad \forall i,j\in V, t\in [n],\\
        &y_i^t \in \{0,1\} \quad \forall i\in V, t\in [n].
    \end{align}

    Observe, that the \textsc{LP} relaxation \textsc{LP-HAC} of the above \textsc{ILP} has an exponential number of constraints, since there can be exponentially many cycles in $G$. However, by \cite{ApproximatingMinimumFeedbackSetsAndMulticutsInDirectedGraphs} we know that there exists a polynomial time separation oracle for the cycle cutting constraints. Therefore, we can solve this \textsc{LP} in polynomial time using the ellipsoid method. Let $LP$ be the solution to \textsc{LP-HAC} and let $c\br{LP}$ be the value of its objective function. We now show how to round the solution of \textsc{LP-HAC} to get a partially integral solution without excessive increase in cost.

    \begin{lemma}\label{lem:lp-rounding}
        Round the values of $y_i^t$ to get $\hat{y}_i^t = 1$ if $y_i^t \geq 1/2$ and $\hat{y}_i^t = 0$ otherwise and $\hat{x}_{ij}^t=\min \brc{1,2y_i^t}$. Then, $\hat{x}, \hat{y}$ is a feasible solution for \textsc{LP-HAC} with cost at most $2\cdot c\br{LP}$.
    \end{lemma}
    \begin{proof}
        Consider any constraint of type (\ref{cycle_constraint_ILP}). If $y_i^t < 1/2$, then $\hat{y}_i^t=0$ and the constraint is satisfied. If $y_i^t \geq 1/2$, then $\hat{y}_i^t=1$. If for any $ij\in C$, $x_{ij}^t=1$ then the constraint is satisfied. Otherwise $\sum_{ij\in C} \hat{x}_{ij}^t = \sum_{ij\in C} 2y_i^t \geq 1$. Consider any constraint of type (\ref{spreading_constraint_ILP}). If $y_i^t \geq 1/2$, then $\hat{y}_i^t=1$ and the constraint is satisfied. If $y_i^t < 1/2$, then $\hat{y}_i^t=0$ and $\sum_{j\in V} \hat{x}_{ij}^t = \sum_{j\in V} 2y_i^t \geq n-t$. The cost of the solution is at most $2\cdot c\br{LP}$ since $\hat{x}_{ij}^t \leq 2x_{ij}^t$ for all $i,j\in V$ and $t\in [n]$.
    \end{proof}

    Fix some choice of $\hat{y}_i^t$ in the above \textsc{LP} so that we have a \textsc{MILP} relaxation \textsc{MILP-HAC} of \textsc{ILP-HAC}. Denote by $MILP$ the solution to \textsc{MILP-HAC} and let $c\br{MILP}$ be the value of its objective function. In what follows, we will need to split this \textsc{MILP} into several instances of \textsc{FES}. We will call such a devided solutions \emph{guides} since they will be used to guide the algorithm about the choice of edges to cut. By $MILP_t^A$ denote the solution of \textsc{MILP-HAC} restricted to the set of vertices $A$ at level $t$ and define $\COST\br{MILP_t^A} = \sum_{ij\in E\br{A}} c\br{i,j} \cdot x_{ij}^t$.

    Now we relax the integrality of $x_{ij}^t$ variables by introducing the $\ell_2^2$-metric constraints. We get the following \textsc{SDP} relaxation \textsc{SDP-MILP-HAC} for the \textsc{MILP-HAC} defined above:
    \begin{align}
        \min \quad &\sum_{t=1}^n \sum_{ij\in E} c\br{i,j} \cdot x_{ij}^t=\sum_{t=1}^n \sum_{ij\in E} c\br{i,j} \cdot\frac{1}{2}\|v_i^t - v_j^t\|_2^2\\
        \text{subject to} \quad &x_{ij}^t \leq x_{ij}^{t-1} \quad \forall i,j\in V, t\in [n],\\
        &x_{ik}^t \leq x_{ij}^t + x_{jk}^t \quad \forall i,j,k\in V, t\in [n],\\
        &x_{ij}^t = \frac{1}{2}\|v_i^t - v_j^t\|_2^2 \quad \forall i,j\in V, t\in [n],\\
        &\|v_i^t\|_2^2 = 1 \quad \forall i\in V, t\in [n],\\
        &\sum_{ij\in C} x_{ij}^t \geq 1 \quad \forall \text{ cycles } C, i\in C, \hat{y}_{i}^t=1, t\in [n],\\
        &\sum_{j\in V} x_{ij}^t \geq n-t \quad \forall i\in V, \hat{y}_{i}^t=0, t\in [n]\\
        &x_{ij}^t \in [0,1] \quad \forall i,j\in V, t\in [n].
    \end{align}

    Denote by $SDP$ the solution to \textsc{SDP-MILP-HAC} and let $c\br{SDP}$ be the value of its objective function. In what follows, we will need to split this \textsc{SDP} into several instances of $k$-\textsc{BP}. By $SDP_t^A$ denote the solution of \textsc{SDP-MILP-HAC} restricted to the set of vertices $A$ at level $t$ and define $\COST\br{SDP_t^A} = \sum_{ij\in E\br{A}} c\br{i,j} \cdot \frac{1}{2}\|v_i^t - v_j^t\|_2^2$. We have the following lemma:

    \begin{lemma}\label{lem:relaxation-decomposition}
        Let $T$ be any HAC tree. For any cluster $A \in \cR^T$, let $r\br{A}$ be the size of the largest child cluster of $A$ if it exists or $0$ otherwise. Fix some $b\in \mathbb{N}^+$. Then:
        \begin{align*}
            \sum_{A\in \cR^T}\sum_{t=r\br{A}/b+1}^{|A|/b} \COST\br{MILP_t^A}\leq c\br{MILP}
            \quad\text{and}\quad
            \sum_{A\in \cR^T}\sum_{t=r\br{A}/b+1}^{|A|/b} \COST\br{SDP_t^A}\leq c\br{SDP}.
        \end{align*}
    \end{lemma}
    \begin{proof}
        It suffices to show that for any fixed edge $ij\in E$ and level $t\in [n]$, the terms $c\br{i,j}\cdot x_{ij}^t$ and $c\br{i,j}\cdot\frac{1}{2}\|v_i^t-v_j^t\|_2^2$, respectively, are each counted at most once in the respective sum. Assume for contradiction that there are two clusters $A$ and $B$ both contributing the term at level $t$. By definition of hierarchical clustering one of $A$, $B$ must be an ancestor of the other; without loss of generality assume $A$ is an ancestor of $B$. Then $t\in \left(r\br{A}/b, |A|/b\right]$ and $t\in \left(r\br{B}/b, |B|/b\right]$, so $bt\in \left(r\br{A}, |A|\right]$ and $bt\in \left(r\br{B}, |B|\right]$. Since $A$ is a proper ancestor of $B$ we have $r\br{A}\geq |B|$, hence $bt\in \left(r\br{A}, |A|\right]\cap\left(r\br{B}, |B|\right]=\emptyset$, a contradiction. The two bounds then follow by summing over all edges and levels.
    \end{proof}

    \begin{lemma}\label{lem:milp-sdp-bound}
        Let $T$ be any HAC tree. Fix some cluster $A \in \cR^T$, some level $t\in [n]$ and $d\in\mathbb{N}^+$, such that $d$ is even. Then, $MILP_t^A/d\leq \sum_{t=\spr{A}/d+1}^{2\spr{A}/d} MILP_t^A$ and $SDP_t^A/d\leq \sum_{t=\spr{A}/d+1}^{2\spr{A}/d} SDP_t^A$.
    \end{lemma}
    \begin{proof}
    Both inequalities are due to the fact that as $t$ increases more edges belong to the cut.
    \end{proof}

\subsection{\textsc{SDP} for \textsc{$k$-BP} and \textsc{LP} for \textsc{FES}}

    By \cite{ApproximatingMinimumFeedbackSetsAndMulticutsInDirectedGraphs} we have the following \textsc{LP} relaxation \textsc{LP-FES} for the \textsc{FES} problem:
    \begin{align}
        \min \quad &\sum_{ij\in E} c\br{i,j} \cdot x_{ij}\\
        \text{subject to} \quad &\sum_{ij\in C} x_{ij} \geq 1, \quad \forall \text{ cycles } C \text{ going through vertex } i\in X,\\
        &0\leq x_{ij} \leq 1, \quad \forall ij\in E.
    \end{align}
    We have the following theorem \cite{ApproximatingMinimumFeedbackSetsAndMulticutsInDirectedGraphs}:
    \begin{theorem}[Even et al. \cite{ApproximatingMinimumFeedbackSetsAndMulticutsInDirectedGraphs}]\label{thm:fes-rounding}
        There exists a polynomial time algorithm which given an instance of the \textsc{FES} problem and a solution to the above \textsc{LP-FES}, finds a subset of edges $S$ such that $G\setminus S$ contains no cycles going through any vertex in $X$ and $\sum_{ij\in S} c\br{i,j} \leq O\br{\log n\log\log n}\cdot \sum_{ij\in E} c\br{i,j} \cdot x_{ij}$.
    \end{theorem}
    The next lemma allows us to decompose the \textsc{MILP-HAC} into several instances of \textsc{LP-FES}:
    \begin{lemma}\label{lem:milp-fes}
        Let $A$ be a non tree cluster of size $r$ and let $r\leq t$ be some fixed level. Then $MILP_r^A$ is a valid solution for the \textsc{FES} problem with the set of vertices $X=\brc{i\in V\br{A}\colon \hat{y}_i^t=1}$.
    \end{lemma}
    \begin{proof}
        In $MILP_r^A$ there are some additional constraints of type $x_{ik}^t\leq x_{ij}^t + x_{jk}^t$ and $\sum_{j\in V} x_{ij}^t \geq n-t$ for every vertex $i$ such that $\hat{y}_i^t=0$. However, this is acceptable since additional constraints yield a stricter relaxation. Now consider any $i\in V$. If $\hat{y}_i^t=0$, then the cycle constraint is always satisfied, which corresponds to the fact that no cycle going through vertex $i$ needs to be cut. If $\hat{y}_i^t=1$, then the cycle constraint is satisfied since it is explicitly included in the \textsc{MILP-HAC} formulation, which corresponds to the fact that every cycle going through vertex $i$ needs to be cut. Moreover, since $i\in X$, we have that all cycle constraints in the \textsc{LP-FES} formulation are included in the \textsc{MILP-HAC} formulation. Therefore, the claim follows.
    \end{proof}
    By \cite{PartitioningGraphsIntoBalancedComponents} we have the following \textsc{SDP} relaxation \textsc{SDP-$k$-BP} for the \textsc{$k$-BP}:
    \begin{align}
        \min \quad &\sum_{ij\in E} c\br{i,j}\cdot \frac{1}{2}\|v_i - v_j\|_2^2\\
        \text{subject to} \quad &\|v_i-v_j\|_2^2+ \|v_j-v_k\|_2^2 \geq \|v_i-v_k\|_2^2, \quad \forall i,j,k\in V,\\
        &\sum_{j\in S} \|v_i-v_j\|_2^2 \geq \spr{S}-\frac{n}{k},\quad \forall S\subseteq V, i\in S.
    \end{align}

    We have the following theorem \cite{PartitioningGraphsIntoBalancedComponents}
    \begin{theorem}[Krauthgamer et al.\cite{PartitioningGraphsIntoBalancedComponents}]\label{thm:kbp-rounding}
        There exists a bicriteria polynomial time algorithm which given an instance of the \textsc{$k$-BP} and a solution to the above \textsc{SDP-$k$-BP}, finds a subset of edges $S$ such that every connected component of $G\setminus S$ has size at most $2n/k$ and $\sum_{ij\in S} c\br{i,j} \leq O\br{\sqrt{\log n}\log k}\cdot \sum_{ij\in E} c\br{i,j}\cdot \frac{1}{2}\|v_i - v_j\|_2^2$.
    \end{theorem}
    The next lemma is a modification of the argument of \cite{ApproximateHierarchicalClusteringViaSparsestCutAndSpreadingMetrics} which shows how to decompose the \textsc{SDP-MILP-HAC} into several instances of the above \textsc{SDP-$k$-BP}:
    \begin{lemma}\label{lem:sdp-kbp}
        Let $A$ be a non tree cluster of size $r$. Then $SDP_t^A$, where $t=r/4$ is a valid solution for \textsc{$k$-BP} with $k=4$.
    \end{lemma}
    \begin{proof}
        In \textsc{SDP-MILP-HAC} at level $t$, there are some additional constraints of type $x_{ij}^t \in [0,1]$, $\|v_i^t\|_2^2=1$ and $\sum_{ij\in C} x_{ij}^t \geq 1$ for every cycle $C$ containing vertex $i$ such that $\hat{y}_i^t=1$. However, this is acceptable since additional constraints yield a stricter relaxation. Now consider the spreading constraints. In \textsc{SDP-MILP-HAC} we have $\sum_{j\in V} x_{ij}^t \geq n-t$ for every vertex $i$ such that $\hat{y}_i^t=0$. Observe that, for no vertex $i\in V\br{A}$ we have $\hat{y}_i^t=1$, therefore, for every $S\subseteq V\br{A}$ we have $\sum_{j\in S} x_{ij}^t\geq \spr{S}-t$ which is the \textsc{SDP-$k$-BP} spreading constraint. Since the triagnle inequality is equivalent in both \textsc{SDP}s the claim follows.
    \end{proof}

    \subsection{The Algorithm}

    We now summarize the algorithm in two parts. The main procedure first performs the preprocessing phase, in which it computes the partially rounded \textsc{LP} solution $(\hat{x},\hat{y})$ and the vector solution $\hat{v}$ obtained from the \textsc{SDP}, and then it invokes the recursive routine. The recursive routine only builds the clustering tree, using the guides prepared by the main procedure.

    \begin{theorem}\label{thm:main-approximation}
        There exists an $O\br{\log n\log\log n}$-approximation algorithm for the \textsc{HAC} problem.
    \end{theorem}
    \begin{proof}

    
    Denote by $\alpha\br{n}$ and $\beta\br{n}$ the approximation ratios of the rounding algorithms for \textsc{FES} and \textsc{$k$-BP} respectively. 

    \input{pseudocodes/main_procedure.tex}
    \input{pseudocodes/recursive_procedure.tex}
    \begin{lemma}\label{lem:recursive-cost}
        Fix sume cluster $H$ for which the recursive procedure was called. Let $A_1^H,\dots, A_{p_H}^H$ be the non-tree children of $H$ of size larger than $|H|/2$. Then, $\COST_G\br{T, u_H}\leq 8\alpha\br{n}\cdot \sum_{t=\spr{H}/8+1}^{\spr{H}} \COST\br{MILP_t^H}$ and $\sum_{i=1}^{p_H}\COST_G\br{T, u_{A_i^H}} \leq 8\beta\br{n}\cdot \sum_{t=\spr{H}/8+1}^{\spr{H}} \COST\br{SDP_t^H}$.
    \end{lemma}
    \begin{proof}
        Firstly, consider the contribution of $u_H$ to the cost. Since it it being cut in a cluster of size $\spr{H}$, we have that $\COST_G\br{T, u_H}=\spr{H}\cdot c\br{F_H}$. By applying Lemma \ref{lem:milp-sdp-bound} with $d=8$ we get that:
        \begin{align*}
            \COST_G\br{T, u_H} = \spr{H}\cdot c\br{F_H} \leq \spr{H}\cdot \COST\br{MILP_{\spr{H}/4}^H} \leq 8\alpha\br{n}\cdot \sum_{t=\spr{H}/8+1}^{\spr{H}/4} \COST\br{MILP_t^H}.
        \end{align*}
        Secondly consider the contribution of all $u_{A_i^H}$ to the cost. Applying Lemma \ref{lem:milp-sdp-bound} again with $d=8$ we get:
        \begin{align*}
             \sum_{i=1}^{p_H}\COST_G\br{T, u_{A_i^H}} &= \sum_{i=1}^{p_H} \spr{A_i^H}\cdot c\br{S_{A_i^H}}\leq \spr{H}\cdot \sum_{i=1}^{p_H} c\br{S_{A_i^H}} \leq \spr{H}\cdot \sum_{i=1}^{p_H} \beta\br{n}\cdot \COST\br{MILP_t^{A_i^H}}\\
             &\leq \spr{H}\cdot \COST\br{MILP_{\spr{H}/4}^H} 
             \leq 8\beta\br{n}\cdot \sum_{t=\spr{H}/8+1}^{\spr{H}/4} \COST\br{SDP_t^H}
        \end{align*}
        where in the first and third inequalities we used the fact that all $A_i^H$ are disjoint subsets of $H$.
    \end{proof}
    Armed with the above lemma we are ready to sum the contribution of all clusters to the cost of the solution. Let $T$ be the clustering tree returned by the algorithm and let $\cR_T$ be the set of all non-tree clusters for which the recursive procedure was called. We have:
    \begin{align*}
        \COST_G\br{T} &= \sum_{H \in \cR_T} \br{\COST_G\br{T, u_H} + \sum_{i=1}^{p_H} \COST_G\br{T, u_{A_i^H}}}
        \\&\leq 8\cdot\sum_{H \in \cH} \sum_{t=\spr{H}/8+1}^{\spr{H}/4} \br{\alpha\br{n}\cdot \COST\br{MILP_t^H} + \beta\br{n}\cdot \COST\br{SDP_t^H}}\\
        &\leq 8\cdot\br{\alpha\br{n}\cdot\COST\br{MILP} + \beta\br{n}\cdot\COST\br{SDP}}\\
        &\leq 16\cdot\br{\alpha\br{n}\cdot\COST\br{LP} + \beta\br{n}\cdot\COST\br{LP}}\\
        &\leq 16\cdot\br{\alpha\br{n}+\beta\br{n}}\cdot \OPT
    \end{align*}
     where the first inequality is due to Lemma \ref{lem:recursive-cost}, the second inequality is due to Lemma \ref{lem:relaxation-decomposition} and the third inequality is due to Lemma \ref{lem:lp-rounding}. 
     
     Since $\alpha\br{n}=O\br{\log n\log\log n}$ and $\beta\br{n}=O\br{\sqrt{\log n}}$, we get an approximation ratio of $O\br{\log n\log\log n}$ for \textsc{HAC}.
    \end{proof}
