\section{General algorithmic framework for approximating \textsc{HC-$\cF$}}

Having characterized the kinds of graph classes that will fit into our framework we can now move on towards describing how to use the properties of $\alpha\br{n}$-well-behavedness to construct a general approximation algorithm for \textsc{HC-$\cF$}. The main idea is that we use a level-based \textsc{LP} relaxation to guide the recursive algorithm.
In the preprocessing step we encode the problem using an \textsc{ILP}. We obtain this formulation by combining any $\alpha\br{n}$-well-behaved \textsc{ILP} with the level decomposition approach from \cite{ApproximateHierarchicalClusteringViaSparsestCutAndSpreadingMetrics}. Our \textsc{ILP} consists of multiple variable levels which encode levels of an $\cF$-clustering tree. It should be noted that these levels are indexed by the size of the clusters, rather than the depth of the tree.

After solving \textsc{LP-HC-$\cF$}, we use a rounding routine to get a partially integral solution which at each level corresponds to choosing which vertices should satisfy the $\cF$-predicate.

During the recursive step, we process clusters in a top-down manner starting from $H=G$. For a cluster $H$, we set $t=w\br{V\br{H}}/2$. Then, we look which vertices satisfy the $\cF$-predicate at level $t$ and which do not (For a formal level definition see Section~\ref{sec:levels} below).
According to this, we partition them into two sets $X_1$ (satisfying the predicate) and $X_0$ (not satisfying the predicate). The solution of \textsc{LP-HC-$\cF$} at level $t$ is then used as a valid solution to \textsc{$p_{\cF}$-P-LP} with $X=X_1$, yielding a cut $F_H$ with $\alpha\br{n}$ factor loss. The same level-$t$ fragment is also used as a valid guide for \textsc{$\rho$-SEP} with $\rho=1/2$, $\rho_0=2/3$ and $X=X_0$, yielding a cut $B_H$, such that each connected component of $H\setminus B_H$ that contains a terminal, has total weight at most $2w\br{V\br{H}}/3$ and the cut cost is within a factor $\beta\br{n}=\bigo\br{\log n}$ of the guide. After cutting the edges in $F_H\cup B_H$, we recursively process all connected components.

Since we decompose the \textsc{LP} into separate levels, by summing up over all of them we will be able to show that our algorithm achieves an approximation ratio of $\bigo\br{\alpha\br{n}+\beta\br{n}}$.

\subsection{Level decomposition of \textsc{HC-$\cF$}} \label{sec:levels}

In order to analyze the structure of $\cF$-clustering trees, we introduce the notion of \emph{level decomposition}, which for a given $\cF$-clustering tree $T$ is defined as follows.
Let $\cC^T$ denote the family of all clusters in $T$, $\cC^T=\brc{G_u\colon u\in V(T)}$.
For $t\in [w\br{G}]$, denote by $\cL_t^T$ the set of all maximal clusters in $\cC^T$ of total weight at most $t$. We call $\cL_t^T$ the $t$-th \emph{level} of the decomposition.
Observe that for each $t$, $\cL_t^T$ is a partition of $V\br{G}$. By $S_t^T$ denote the set of edges with endpoints in different clusters of $\cL_t^T$. We call $S_t^T$ the $t$-th \emph{level cut}.
It is easy to see that the cost of $T$ can be expressed as:
\begin{observation}\label{obs:cost-decomposition}
    $\COST_G\br{T} = \sum_{t=1}^{w\br{G}} c\br{S_t^T}$.
\end{observation}

Since we use the above levels to decompose an $\cF$-clustering tree and its cost, we would also like to do the same for the $\cF$-predicate. This is because eventually, all vertices satisfy the $\cF$-predicate, but whether a vertex $v\in V$ satisfies the $\cF$-predicate, when it belongs to a cluster at the $t$-th level depends on the structure of the $\cF$-clustering tree.
For every level $t\in [w\br{G}]$ and every vertex $v\in V\br{G}$, we define the level-indexed $\cF$-predicate indicator by setting $p_{\cF}^t\br{v}$ to be true iff $p_{\cF}\br{v}$ is true in $G\setminus S_t^T$. 

\subsection{\textsc{ILP} formulation of \textsc{HC-$\cF$} and its relaxation}
    We now move towards a description of the constraints of the \textsc{ILP} along with an explanation of the role they play. We use the well-known \emph{spreading-metrics} framework which in recent years have come particularly useful for approximating many problems regarding graph cutting \cite{ExpanderFlowsGeometricEmbeddingsAndGraphPartitioning,PartitioningGraphsIntoBalancedComponents,ApproximateHierarchicalClusteringViaSparsestCutAndSpreadingMetrics,FastApproximateGraphPartitioningAlgorithms,DivideAndConquerApproximationAlgorithmsViaSpreadingMetrics}.
    Spreading metrics are a way to encode the structure of a graph into a metric space, which can then be used to find cuts in the graph. The task of a mathematical program (usually LP or SDP) is to find a metric that spreads out the vertices of the graph in a way that reflects the structure of the graph. Here, the idea is to merge such constraints with our level-indexed formulation to encode \textsc{HC-$\cF$} as a linear program.

    % By our definition, for every level $t\in [w\br{G}]$, each cluster is either of total weight at most $t$ or it is an $\cF$-cluster. Moreover, we observe, that the partition at level $t-1$ refines the partition at level $t$.
    For each edge $e\in E$ and level $t\in [w\br{G}]$ we use a variable $y_e^t\in[0,1]$ which encodes whether $e\in S_t^T$.
    We also define a variable $x_v^t=1$ iff $p_{\cF}^t\br{v}$ is true and $0$ otherwise.  Let $\dist_{y, V}^{t}\colon V\times V\to \mathbb{R}_{\geq 0}$ denote shortest-path distances in $G$ induced by lengths $\{y_e^t\}_{e\in E}$.
    We first explain all constraints and then we give the full ILP. 
    These are as follows:
    \begin{enumerate}
        \item \textbf{refinement:} $y_{e}^t \leq y_{e}^{t-1}$ for all $e\in E$ and $t\in [w\br{G}]$, because the partition at level $t-1$ refines the partition at level $t$. Moreover, $x_v^t \leq x_v^{t-1}$ for all $v\in V$ and $t\in [w\br{G}]$, since $p_{\cF}^t\br{v}$ implies $p_{\cF}^{t-1}\br{v}$.

        \item \textbf{$p_{\cF}$-P constraints:} Recall, that for every vertex $v\in V$, the set $\cR_v$ contains all subsets of edges $R\subseteq E$ such that if at least one edge in each $R\in \cR_v$ is cut, then $p_{\cF}\br{v}$ is true. Therefore, for each $v\in V$, every level $t\in [w\br{G}]$ and every $R\in \cR_v$ we have $\sum_{e\in R} y_e^t \geq x_{v}^t$. Additionally, we have $x_v^1=1$, since $p_{\cF}^1\br{v}$ is true for all $v\in V$.

        \item \textbf{spreading-metrics:} For every vertex $v\in V$, every level $t\in [w\br{G}]$, and every subset $S\subseteq V$ containing $v$, we have
        \[
            \sum_{u\in S} w(u)\cdot \dist_{y, V}^{t}(v,u) \geq \br{1-x_v^t}\cdot \br{w\br{S}-t}.
        \]
        When $x_v^t=0$, this condition enforces that inside any set $S$ containing $v$, enough weighted mass is separated from $v$ unless the cluster weight around $v$ is already at most $t$. When $x_v^t=1$, the right-hand side vanishes, corresponding to the fact that the cluster containing $v$ is an $\cF$-cluster and hence its weight can be larger than $t$.
    \end{enumerate}

    The following \textsc{ILP}, which we call \textsc{ILP-HC-$\cF$}, is an exact formulation of the \textsc{HC-$\cF$} problem:
    \begin{align}
        \min \quad &\sum_{t=1}^{w\br{G}} \sum_{e\in E} c(e) \cdot y_{e}^t\\
        \text{subject to} \quad &y_{e}^t \leq y_{e}^{t-1} \quad \forall e\in E,\; t\in [w\br{G}],\\
        &x_v^t \leq x_v^{t-1} \quad \forall v\in V, t\in [w\br{G}],\\
        &x_v^1 = 1 \quad \forall v\in V,\\
        &\sum_{e\in R} y_{e}^t \geq x_{v}^t \quad \forall v\in V, R\in \cR_v, t\in [w\br{G}], \label{cover_constraint_ILP}\\
        &\sum_{u\in S} w(u)\cdot \dist_{y, V}^{t}(v,u) \geq \br{1-x_v^t}\cdot \br{w\br{S}-t} \quad \forall S\subseteq V, v\in S, t\in [w\br{G}], \label{spreading_constraint_ILP}\\
        &y_{e}^t \in \{0,1\} \quad \forall e\in E, t\in [w\br{G}],\\
        &x_v^t \in \{0,1\} \quad \forall v\in V, t\in [w\br{G}].
    \end{align}
    We denote by $\OPT_{ILP}$ the value of an optimal solution to the above ILP for a given graph $G$.

    \begin{lemma}\label{lem:ilp-exact}
        \textsc{ILP-HC-$\cF$} is an exact formulation of \textsc{HC-$\cF$} which implies that $\OPT_{ILP}=\OPT_{\cF}$.
    \end{lemma}
    \begin{proof}
        $(\Rightarrow)$ Let $T$ be any feasible $\cF$-clustering tree.
        Recall that, for each level $t\in [w\br{G}]$, $S_t^T$ is the level cut from the level decomposition and set
        \[
            y_e^t:=\mathbf{1}[e\in S_t^T].
        \]
        Also set $x_v^t:=1$ if and only if the level-$t$ cluster of $v$ satisfies $p_{\cF}$.

        The refinement constraints hold because level cuts are nested. The constraints on $x$ hold by monotonicity of the predicate along the refinement. The $p_{\cF}$-P constraints hold by definition of the family $\cR_v$: whenever $x_v^t=1$, each required $p_{\cF}$-P constraint for $v$ at level $t$ is satisfied by the cut pattern encoded by $y^t$.

        For spreading constraints, fix $v\in V$, $t\in [w\br{G}]$, and $S\subseteq V$ with $v\in S$. If $x_v^t=1$, the right-hand side is $0$. Assume $x_v^t=0$ and let $C$ be the level-$t$ cluster containing $v$. Then $C$ is not an $\cF$-cluster, so by level definition $w(C)\le t$. Every $u\in S\setminus C$ is separated from $v$ by level-$t$ cuts, hence $\dist_{y, V}^{t}(v,u)\ge 1$. Therefore,
        \[
            \sum_{u\in S} w(u)\cdot \dist_{y,V}^{t}(v,u)
            \ge \sum_{u\in S\setminus C} w(u)
            = w(S)-w(S\cap C)
            \ge w(S)-w(C)
            \ge w(S)-t,
        \]
        which is exactly \eqref{spreading_constraint_ILP} for $x_v^t=0$. Thus $(y,x)$ is ILP-feasible and
        \[
            \sum_{t=1}^{w\br{G}}\sum_{e\in E} c(e)\cdot y_e^t
            = \sum_{t=1}^{w\br{G}} c(S_t^T)
            = \COST_G(T)
        \]
        by Observation~\ref{obs:cost-decomposition}. Hence $\OPT_{ILP}\le \OPT_{\cF}$.

        $(\Leftarrow)$ Let $(y,x)$ be any integral feasible ILP solution. For each level $t$, define
        \[
            E_t^0:=\brc{e\in E\colon y_e^t=0},
        \]
        and let $\Pi_t$ be the partition of $V$ into connected components of $G\setminus \br{E\setminus E_t^0}$. (Note that $E\setminus E_t^0$ are exactly the edges that are considered to be cut at level $t$). Because $y_e^t\le y_e^{t-1}$, we have $E_{t-1}^0\subseteq E_t^0$, so $\Pi_{t-1}$ refines $\Pi_t$ and the family $(\Pi_t)_{t=1}^{w\br{G}}$ is laminar; therefore it defines a $\cF$-clustering tree $T$.

        Consider any level $t$ and any component $C\in \Pi_t$. For every $v\in C$, all vertices of $C$ are at zero $\dist_{y,V}^t$-distance from $v$, so plugging $S=C$ into \eqref{spreading_constraint_ILP} gives
        \[
            0\ge (1-x_v^t)\cdot (w(C)-t).
        \]
        Hence either $w(C)\le t$, or $x_v^t=1$ for all $v\in C$. In the second case, by $p_{\cF}$-P constraints and the definition of $\cR_v$, each vertex in $C$ satisfies the $\cF$-predicate; therefore $C\in \cF$. So every cluster in level $t$ is either of weight at most $t$ or belongs to $\cF$, exactly as required in a $\cF$-clustering tree.

        Finally, the level cut induced by $\Pi_t$ is precisely $\brc{e\in E\colon y_e^t=1}$, so
        \[
            \COST_G(T)=\sum_{t=1}^{w\br{G}} c\br{\brc{e\in E\colon y_e^t=1}}
            =\sum_{t=1}^{w\br{G}}\sum_{e\in E} c(e)\cdot y_e^t.
        \]
        Therefore every integral ILP solution maps to a feasible $\cF$-clustering tree of the same cost, implying $\OPT_{\cF}\le \OPT_{ILP}$, which completes the proof.
    \end{proof}

    We now wish to show that it is possible to solve the relaxation of \textsc{ILP-HC-$\cF$} in polynomial time. We denote this relaxation by \textsc{LP-HC-$\cF$}, which is obtained by replacing the integrality constraints on $x_v^t$ and $y_e^t$ with $x_v^t\in [0,1]$ and $y_e^t\geq 0$ for all $v\in V$, $e\in E$, and $t\in [w\br{G}]$. The following lemma shows that \textsc{LP-HC-$\cF$} can be solved in polynomial time.
    \begin{lemma}\label{lem:lp-polytime}
        The relaxation \textsc{LP-HC-$\cF$} can be solved in polynomial time.
    \end{lemma}
    \begin{proof}
        The only potentially exponential families of constraints are \eqref{cover_constraint_ILP} and \eqref{spreading_constraint_ILP}. For constraints of type \eqref{cover_constraint_ILP}, a polynomial-time separation oracle is part of the assumptions of a well-behaved formulation.

        It remains to separate the spreading-metrics constraints. Fix a level $t\in [w\br{G}]$ and a vertex $v\in V$. Using the current values of the variables $y_e^t$, we compute all distances $\dist_{y,V}^{t}(v,u)$ for $u\in V$ by a shortest-path algorithm. Then the left-hand side minus the right-hand side of \eqref{spreading_constraint_ILP} can be rewritten as
        \[
            \sum_{u\in S} w(u)\cdot \dist_{y,V}^{t}(v,u) - \br{1-x_v^t}\cdot \br{w\br{S}-t}
            = \br{1-x_v^t}\cdot t - \sum_{u\in S} w(u)\cdot \br{\br{1-x_v^t}-\dist_{y,V}^{t}(v,u)}.
        \]
        Therefore, for fixed $v$ and $t$, the most violated constraint is obtained by taking
        \[
            S=\brc{u\in V\colon \dist_{y,V}^{t}(v,u)<1-x_v^t},
        \]
        since every vertex with positive contribution to the second term should be included, while vertices with non-positive contribution should be excluded. Hence, after one shortest-path computation, we can test in polynomial time whether any spreading-metrics constraint for this pair $(v,t)$ is violated and, if so, output a violated one.

        Repeating this for all polynomially many pairs $(v,t)$ yields a polynomial-time separation oracle for constraints of type \eqref{spreading_constraint_ILP}. Consequently, \textsc{LP-HC-$\cF$} admits a polynomial-time separation oracle, and thus it can be solved in polynomial time by the ellipsoid method.
    \end{proof}
    %Note that in our relaxation we allow each $y_e^t\geq 0$ and each $x_v^t$ to take any value in $[0,1]$.
    Let $LP$ be a solution to \textsc{LP-HC-$\cF$} and let $\OPT_{LP}$ be the optimal value of its objective function. We now show how to round the solution of \textsc{LP-HC-$\cF$} to get a partially integral solution without excessive increase in cost.

    \begin{lemma}\label{lem:lp-rounding}
        Round the values of $x_v^t$ to get $\hat{x}_v^t = 1$ if $x_v^t \geq 1/2$ and $\hat{x}_v^t = 0$ otherwise
        and $\hat{y}_{e}^t=2\cdot y_{e}^t$. Then, $\br{\hat{y}, \hat{x}}$ is a feasible solution for \textsc{LP-HC-$\cF$} with cost at most $2\cdot \OPT_{LP}$.
    \end{lemma}
    \begin{proof}
        Firstly, observe that the refinement constraints are satisfied since $\hat{y}_{e}^t \leq \hat{y}_{e}^{t-1}$ and $\hat{x}_v^t \leq \hat{x}_v^{t-1}$ for all $e\in E$, $v\in V$ and $t\in [w\br{G}]$. The $p_{\cF}$-P constraints are satisfied since if $\hat{x}_v^t=1$, then $x_v^t\geq 1/2$ and hence $\sum_{e\in R} y_{e}^t \geq x_v^t \geq 1/2$ for all $R\in \cR_v$. Therefore, $\sum_{e\in R} \hat{y}_{e}^t = 2\cdot \sum_{e\in R} y_{e}^t \geq 1$. If $\hat{x}_v^t=0$, then the $p_{\cF}$-P constraint is trivially satisfied. The spreading-metrics constraints are satisfied since if $\hat{x}_v^t=0$, then $x_v^t<1/2$ and hence
        \[
            \sum_{u\in S} w(u)\cdot \dist_{y,V}^{t}(v,u) \geq \br{1-x_v^t}\cdot \br{w\br{S}-t} > 1/2\cdot \br{w\br{S}-t}.
        \]
        Therefore, we get that for all $S\subseteq V$ containing $v$, such that $\hat{x}_v^t=0$, we have
        \[
            \sum_{u\in S} w(u)\cdot \dist_{\hat{y},V}^{t}(v,u) = 2\cdot \sum_{u\in S} w(u)\cdot \dist_{y,V}^{t}(v,u) > w\br{S}-t
        \]
        as required.
        If $\hat{x}_v^t=1$, then the spreading-metrics constraint is trivially satisfied.
        Finally, the cost of the new solution is at most $2\cdot \OPT_{LP}$ since $\sum_{t=1}^{w\br{G}} \sum_{e\in E} c(e) \cdot \hat{y}_{e}^t = 2\cdot \sum_{t=1}^{w\br{G}} \sum_{e\in E} c(e) \cdot y_{e}^t = 2\cdot \OPT_{LP}$.
    \end{proof}

    Overall, we obtain a partially integral point where all of the $x$ values are integral, which translates to the fact that we have decided which vertices need to satisfy the $\cF$-predicate at each level of the $\cF$-clustering tree. From now on we will thus assume that for every $v\in V$ and $t\in[w\br{G}]$, the value of $p_{\cF}^t\br{v}$ is fixed and we can remove all of the $x$ variables from the \textsc{LP}. Since we will not use the original values in further considerations, by a slight abuse of notation from now on we will refer to this solution simply by $LP$.

\subsection{LP relaxation for $\rho$-separators}\label{sec:lp-rho-sep}

    We now state the linear programming relaxation for the \textsc{$\rho$-SEP} which is a modified LP of \cite{FastApproximateGraphPartitioningAlgorithms} that will be required in order to reduce the size of the instance efficiently. For the \textsc{$\rho$-SEP} problem on a graph $G=(V,E)$ with a terminal set $X\subseteq V$, we use the following weighted spreading-metrics \textsc{LP} relaxation \textsc{LP-$\rho$-SEP}:
    \begin{align}
        \min \quad &\sum_{e\in E} c(e)\cdot y_e\\
        \text{subject to} \quad &\sum_{u\in S} \dist_{y,V}(v,u)\cdot w(u) \geq w\br{S}-\rho\cdot w\br{G}, \quad \forall S\subseteq V,\; v\in S\cap X,\\
        &0\leq y_e\quad \forall e\in E.
    \end{align}

    The analysis in \cite{FastApproximateGraphPartitioningAlgorithms} is stated for the special case $X=V$. In our framework, we use the corresponding extension for the case $X\neq V$: the same sphere-growing argument is applied, the only important modification is that we grow the spheres around centers in $X$, rather than around arbitrary vertices in $V$. We state the resulting guarantee below and defer the full extension from $X=V$ to $X\neq V$ to Appendix~\ref{app:terminal-extension}.

    The following bicriteria rounding guarantee for \textsc{$\rho$-SEP} is a direct corollary of \cite{FastApproximateGraphPartitioningAlgorithms}:
    \begin{theorem}\label{thm:ksep-rounding}
        There exists a polynomial-time algorithm which, given an instance of \textsc{$\rho$-SEP} and a solution to the above \textsc{LP-$\rho$-SEP}, returns a subset of edges $S$ such that every connected component of $G\setminus S$ that contains a terminal has total weight at most $\rho_0\cdot w\br{G}$ and
        \[
            \sum_{ij\in S} c\br{i,j} \leq \bigo\br{\log n}\cdot \sum_{e\in E} c(e)\cdot y_e,
        \]
        for any fixed constants $0<\rho<\rho_0<1$.
    \end{theorem}

\subsection{Restricting the solutions}
    
In order to obtain quality approximation ratios we employ a charging scheme, where we split the $LP$ into several solutions restricted to the clusters of the $\cF$-clustering tree $T$ we are building. Fix some cluster $H$ of $G$ and a level $t$. By $LP_t^H$ denote the solution of \textsc{LP-HC-$\cF$} restricted to the set of vertices $H$ at level $t$ which is done by discarding all of the variables $y_{uv}^t$, such that $\brc{u,v}\cap V\br{H}\neq \brc{u,v}$ or levels other than $t$. We call each such restricted part $LP_t^H$ a \emph{guide}. For any target problem $\Pi$ with LP relaxation \textsc{LP-$\Pi$} (on the same graph fragment and terminal set), we say that $LP_t^H$ is a \emph{valid guide} for $\Pi$ if the restricted values from $LP_t^H$ satisfy all constraints of \textsc{LP-$\Pi$}. In other words, a valid guide is simply a feasible LP point for the target problem, inherited from \textsc{LP-HC-$\cF$}. This restriction is also performed on the cost function and this is reflected in the accordingly defined quantities $\COST\br{LP_t^H} = \sum_{uv\in E\br{H}} c\br{u,v} \cdot y_{uv}^t$. We have the following lemma:

    \begin{lemma}\label{lem:relaxation-decomposition}
        Let $T$ be any $\cF$-clustering tree. For any cluster $H \in \cC^T$, let $r\br{H}$ be the weight of the heaviest child cluster of $H$ if it exists or $0$ otherwise. Moreover, fix some constant $g\in \mathbb{N}$. Then:
        \begin{align*}
            \sum_{H\in \cC^T}\sum_{t=r\br{H}+1}^{w\br{V\br{H}}} c\br{LP_{\fl{t/g}}^H}\leq g\cdot \OPT_{LP}
        \end{align*}
    \end{lemma}
    \begin{proof}
        For convenience, define $y_{u}^0=y_{u}^1$ for all $u\in V$. The value $c\br{LP_{0}}$ is defined analogously as before. 
        We start by showing that 
        \[
        \sum_{H\in \cC^T}\sum_{t=r\br{H}+1}^{w\br{V\br{H}}} c\br{LP_{\fl{t/g}}^H}\leq \sum_{i=0}^{w\br{G}} c\br{LP_{\fl{i/g}}^G}.
        \]
        It suffices to show that for any fixed edge $uv\in E$ and level $t\in [w\br{G}]$, the term $c\br{u,v}\cdot y_{uv}^t$ is counted at most once in the above sum. Assume for contradiction that there are two clusters $H$ and $B$ both contributing this term at level $t$. By definition of hierarchical clustering, one of $H$, $B$ must be an ancestor of the other; without loss of generality assume $H$ is an ancestor of $B$. Then
        \[
            t \in \left(r\br{H}, w\br{V\br{H}}\right]\quad \text{and} \quad
            t \in \left(r\br{B}, w\br{V\br{B}}\right].
        \]
        Since $H$ is an ancestor of $B$, we have $r\br{H} \geq w\br{V\br{B}}.$
        Hence the above intervals are disjoint, a contradiction. The bound follows by summing over all edges and levels.

        Now, observe that by simple reindexing, we have that 
        \[\sum_{i=0}^{w\br{G}} c\br{LP_{\fl{i/g}}^G}\leq g\cdot\sum_{i=1}^{w\br{G}} c\br{LP_{i}^G}.\]
        But the right-hand side is exactly $g\cdot \OPT_{LP}$, which completes the proof.
    \end{proof}

    \begin{lemma}\label{lem:lp-level-bound}
        Let $H$ be any cluster and let $d, b\in\mathbb{N}^+$. Then
        \[
        w\br{V\br{H}}\cdot c\br{LP_{\fl{w\br{V\br{H}}}/b}^H}\leq d\cdot \sum_{t=\br{1-1/d}\cdot w\br{V\br{H}}+1}^{w\br{V\br{H}}} c\br{LP_{\fl{t/b}}^H}.
        \]
    \end{lemma}
    \begin{proof}
    The inequality follows from monotonicity of level cuts: as $t$ decreases, more edges belong to the cut.
    \end{proof}

    Fix a subgraph $H$ of $G$ with weight $r=w\br{V\br{H}}$ and a level $t\in [r]$. Define
    \[
        X_1:=\brc{v\in V\br{H}\colon \hat{x}_v^t=1}, \qquad X_0:=\brc{v\in V\br{H}\colon \hat{x}_v^t=0}.
    \]

    The next lemma allows us to decompose a solution to \textsc{LP-HC-$\cF$} into several instances of \textsc{LP-$p_{\cF}$-P}:
    \begin{lemma}\label{lem:lp-fes}
        Let $H$ be cluster that does not belong to $\cF$, with weight $r=w\br{V\br{H}}$ and let $t\leq r$ be some fixed level. Then $LP_r^H$ is a valid solution for the \textsc{$p_{\cF}$-P} for $X=X_1$.
    \end{lemma}
    \begin{proof}
        In $LP_r^H$ there are additional metric/spreading constraints for vertices $v$ such that $p_{\cF}^t(v)=0$. This is acceptable since additional constraints yield a stricter relaxation. Now consider any $v\in V$. If $p_{\cF}^t(v)=0$, then the $p_{\cF}$-P constraint is always satisfied, which corresponds to the fact that $v$ does not need to belong to a $\cF$-cluster. If $p_{\cF}^t(v)=1$, then the $p_{\cF}$-P constraint is satisfied since it is explicitly included in the \textsc{LP-HC-$\cF$} formulation. Moreover, since $v\in X_1$, we have that all $p_{\cF}$-P constraints in the \textsc{LP-$p_{\cF}$-P} formulation are included in the \textsc{LP-HC-$\cF$} formulation. Therefore, the claim follows.
    \end{proof}
    \begin{lemma}\label{lem:lp-ksep}
        Let $H$ be a non $\cF$-cluster of weight $r=w\br{V\br{H}}$ and let $t<r$. Then $LP_t^H$ is a valid guide for the weighted spreading-metrics \textsc{LP-$\rho$-SEP} with $\rho=t/r$ for $X=X_0$.
    \end{lemma}
    \begin{proof}
        The statement is analogous to Lemma~\ref{lem:lp-fes}, but with the complement terminal set.
        First, the $p_{\cF}$-P constraints are irrelevant for \textsc{$\rho$-SEP} feasibility and can only make the relaxation stronger.

        Second, consider the vertices in $X_0$ (i.e., those with $p_{\cF}^t(v)=0$). For every such $v$ and every $S\subseteq V(H)$ with $v\in S$, the spreading-metrics constraint of \textsc{LP-HC-$\cF$} at level $t$ gives
        \[
            \sum_{u\in S} w(u)\cdot \dist_{y,V}^{t}(v,u) \geq w\br{S}-t,
        \]
        where $\dist_{y,V}^{t}$ denotes shortest-path distances taken in the whole graph $G$.

        On the other hand, in \textsc{LP-$\rho$-SEP} for the cluster $H$, the relevant distances are the ones measured inside $H$, namely $\dist_{y,V(H)}^{t}$. Since $H$ is a subgraph of $G$, every $u$-$v$ path contained in $H$ is also a path in $G$, whereas paths leaving $H$ are no longer available; restricting the graph to the cluster $H$ can therefore only remove paths and never shorten the distances. Consequently,
        \[
            \dist_{y,V(H)}^{t}(v,u) \geq \dist_{y,V}^{t}(v,u) \qquad \text{for all } u,v\in V(H).
        \]
        Combining the two inequalities, for every $S\subseteq V(H)$ and every $v\in S\cap X_0$ we obtain
        \[
            \sum_{u\in S} w(u)\cdot \dist_{y,V(H)}^{t}(v,u) \geq \sum_{u\in S} w(u)\cdot \dist_{y,V}^{t}(v,u) \geq w\br{S}-t.
        \]
        With $t$ and $V\br{H}$ as the metric, this is exactly the weighted spreading requirement of \textsc{LP-$\rho$-SEP} corresponding to $\rho=t/r$. Hence, all constraints of \textsc{LP-$\rho$-SEP} are satisfied, and $LP_t^H$ is a feasible \textsc{LP-$\rho$-SEP} guide for $X_0$.
    \end{proof}
