\section{Preliminaries}

Throughout this paper, each considered set of points along the similarity function is treated as an undirected weighted graph $G=\br{V,E,c,w}$, where $c\colon E\to \mathbb{N}$ is the similarity function (if $uv\notin E$, we may assume as a convention that $c(uv)=0$) and $w\colon V\to \mathbb{N}^+$ are vertex weights. We write $n\br{G}=\spr{V}$, $m\br{G}=\spr{E}$, and $n=n\br{G}$, $m=m\br{G}$ when $G$ is clear from context. For $U\subseteq V$, let $w\br{U}=\sum_{v\in U} w(v)$ and let $w\br{G}=w\br{V}$. For a set of edges $F\subseteq E$, we use $c\br{F}=\sum_{e\in F} c(e)$. Let $[n]=\brc{1,\dots, n}$.
By a permutation $\pi$ of $V$ we mean a bijection $\pi\colon V\to [n]$.

Consider a tree $T$ constructed for a graph $G$ in which each node $u$ corresponds to a \emph{cluster} $G_u$ (an induced subgraph of $G$), the root corresponds to $G$, and the children of a node $u$ form a partition of $G_u$ into connected components.
The set of edges of $G_u$ having endpoints in different connected components corresponding to the children of $u$ are denoted by $S_u$.
Depending on the problem we either require that the connected components corresponding to the leaves of $T$ are singletons in which case we refer to $T$ as a \emph{clustering tree}, or belong to the desired class $\cF$ and then $T$ is called a \emph{$\cF$-clustering tree}.

\subsection{Problem definitions}\label{subsec:problems-definitions}

For a given family of graphs $\cF$, let $p_{\cF}\colon V\to \brc{0,1}$ be a predicate function, which to each vertex assigns a Boolean value encoding whether a given vertex satisfies some structural property. We say that $p_{\cF}$ is an \emph{$\cF$-predicate} if for any graph $G$, we have that for every connected component $H$ of $G$, $H\in \cF$ whenever $\bigwedge_{v \in V}p_{\cF}\br{v}$ is true. Although this definition might seem a bit abstract, in what follows we show that the graph classes of our interest have naturally defined $\cF$-predicates. It should also be noted that for any graph class there might be many different $\cF$-predicates, but for our needs it is sufficient to consider only one of them per class.
Given a weighted graph $G=\br{V, E, c, w}$, we define the following combinatorial problems:
\begin{itemize}
    \item \textsc{Hierarchical Clustering} (\textsc{HC}): find a clustering tree $T$ of $G$ that minimizes $\COST_G\br{T} = \sum_{u\in V\br{T}} c(S_u) \cdot w\br{V\br{G_u}}$.
    \item \textsc{Hierarchical $\cF$-Clustering} (\textsc{HC-$\cF$}): find a $\cF$-clustering tree $T$ of $G$ that minimizes $\COST_G\br{T} = \sum_{u\in V\br{T}} c(S_u) \cdot w\br{V\br{G_u}}$.
    \item \textsc{$p_{\cF}$-Partitioning} (\textsc{$p_{\cF}$-P})\label{prob:pfp}:
    Given an additional set $X\subseteq V$ of terminals, find a subset of edges $E'\subseteq E$ such that for each $v\in X$, the $\cF$-predicate $p_{\cF}\br{v}$ is satisfied in $G\setminus E'$ and the cost $c\br{E'}$ is minimized.
    The following problems are special cases of \textsc{$p_{\cF}$-P}:
    \begin{enumerate}
        \item \textsc{Feedback Edge Set} (\textsc{FES}): given an additional set $X$ of vertices, find a subset of edges $E'\subseteq E$ such that $G\setminus E'$ contains no cycles going through vertices in $X$ and $c\br{E'}$ is minimized. Here, the $\cF$-predicate is simply whether there exists a cycle in $G\setminus E'$ going through $v$.
    \item \textsc{Min-Multicut} (\textsc{MMC}): given a set of vertex pairs $\{(s_i, t_i)\in V\times V\colon i\in [k]\}$, find a subset of edges $E'\subseteq E$ such that each pair $(s_i, t_i)$ is disconnected in $G\setminus E'$ and $c\br{E'}$ is minimized. The problem of partitioning into $\cD_d$-clusters of fixed diameter $d$ can be reduced to \textsc{MMC} by setting the terminal pairs to be all $X \times V$ pairs such that $\dist\br{u, v} > d$, since such a pair cannot belong to the same connected component of $G\setminus E'$\footnote{The distance $\dist$ can be defined with respect to either number of edges, $c$ or any other metric.}. Here for a given vertex $v\in V$, the $\cF$-predicate is whether in the connected component $H\in G\setminus E'$ containing $v$, there exists a vertex $u$ such that $\dist\br{u, v} > d$.
    \item \textsc{$\rho$-Separator} (\textsc{$\rho$-SEP})\label{prob:rho-sep}: given a parameter $\rho\in(0,1)$ and an additional set $X\subseteq V$ of vertices, find a subset of edges $E'\subseteq E$ such that every connected component $H$ of $G\setminus E'$ with $V(H)\cap X\neq\emptyset$ has the total vertex weight at most $\rho\cdot w\br{G}$ and $c\br{E'}$ is minimized. We call such an $E'$ a $(\rho,X)$-separator. Here, the $\cF$-predicate is whether the connected component $H\in G\setminus E'$ containing $v$ has the total vertex weight at most $\rho\cdot w\br{G}$. A bicriteria approximation algorithm for \textsc{$\rho$-SEP} with parameters $(\rho,\rho_0)$, where $0<\rho<\rho_0<1$, returns a subset of edges $E'$ such that $c\br{E'}$ is at most $\rho$ times more then the optimum and every connected component of $H\in G\setminus E'$ with $H\cap X\neq\emptyset$ has total vertex weight of at most $\rho_0\cdot w\br{G}$.
    \end{enumerate}
    \item \textsc{Minimum Linear Arrangement} (\textsc{MLA}): find a permutation $\pi$ of the vertices of $G$ that minimizes $\COST_G\br{\pi} = \sum_{uv\in E} c(uv) \cdot \spr{\pi(u) - \pi(v)}$.
\end{itemize}

The inclusion of weights serves the purpose of obtaining a more general solution, however its occurrence does not impact the techniques nor the results in a significant way.
On the other hand, we point out that our working assumption will be that the value $w\br{G}$ is polynomial in $n$.
If this assumption is not met for a given input, then the algorithms work correctly but become pseudo-polynomial in the size of the graph.

For any node $u\in V(T)$ of a clustering (or $\cF$-clustering) tree $T$, we denote by $\COST_G\br{T, u} = c(S_u) \cdot w\br{V\br{G_u}}$ the contribution of the node $u$ to the total cost of the tree.
By $\OPT_{\cF}$ we denote the cost of an optimal solution to \textsc{HC-$\cF$}.

We further remark, that the introduction of $\cF$-predicates allows us to use somehow weaker conditions to encode progress of our algorithm while trying to find a solution to \textsc{HC-$\cF$}. In particular, consider the case $\cF=\cT$. Here, there are at least two natural choices for $p_{\cF}$. For our needs, we want that there is no cycle in $G$ going through a given vertex $v\in V$. Observe however, that we also could define the $p_{\cF}$ to encode whether the cluster containing $v$ is a tree. The latter condition is stronger, but the former one allows us to use already existing approximation algorithms for \textsc{FES}. As it turns out, either way our ILP formulation will be forced to cut edges of a given cluster if at least one vertex of this cluster does not satisfy the $\cF$-predicate. Therefore, we can use the weaker condition to encode progress of our procedure which greatly eases the design of the algorithm. This relaxed condition also potentially allows for finding more classes of graphs which fit our framework.
